---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
        activity <- read.csv(unzip("./activity.zip"))
        head(activity)
```
## What is mean total number of steps taken per day?
```{r echo=TRUE, results='markup'}
        library(dplyr)
        grouped_activity <- activity %>% 
                group_by(date) %>% 
                summarise(total_steps = sum(steps))
        hist(grouped_activity$total_steps)
        mean_total_steps <- mean(grouped_activity$total_steps, na.rm = TRUE)
        median_total_steps <- median(grouped_activity$total_steps, na.rm = TRUE)
```
The mean total number of steps taken per day is **`r mean_total_steps`**

The median total number of steps taken per day is **`r median_total_steps`**


## What is the average daily activity pattern?

```{r}
        library(dplyr)
        avgstep <- activity %>% 
                group_by(interval) %>% 
                summarise(avg_step = mean(steps, na.rm = TRUE))
        plot(x = avgstep$interval, y = avgstep$avg_step, type = "l")
        max_steps <- max(avgstep$avg_step)
        max_interval <- avgstep[avgstep$avg_step == max_steps,]$interval
```


Interval number `r max_interval` contains the highest average step count of `r max_steps`


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
